Summer 2025 ToDo

### 2025-10-01 Relations/Paper/Well-foundedness.tex
WIP swapping control
well-foundedness

On the other hand, assuming accessibility is $\lnot\lnot$-closed, the other
hypothesis can now be weakened, to only require existence of minimal elements
for $\lnot\lnot$-closed subsets.  This is a significant restriction, which may
be validated in many cases of interests.  In contrast, the original hypothesis
(postulating minimal elements for ALL predicates)
is generally equivalent to every subset being decidable.
For example, the file \texttt{WFCounters.agda} shows that assuming the full minimality hypothesis
for the standard strict order on the natural numbers entails decidability
of every subset of $\nat$ --- which is clearly incompatible with effective semantics.
\apcomment{Make sure the file exists (Done), check it wrt other new stuff in the FPLA repo.}

One could also consider proving well-foundedness ``up to $\lnot\lnot$'',
asserting that every element is $\lnot\lnot$-accessible,
every sequence $\lnot\lnot$-contains a non-decreasing index,
for every non-empty subset there $\lnot\lnot$-exists a minimal element, etc.
These are formalized in the definition below, and in the corresponding file
\texttt{WFWeakDefinitions.agda}.

@@

>>>>

On the other hand, assuming accessibility is $\lnot\lnot$-closed, the other
hypothesis can now be weakened, to only require existence of minimal elements
for $\lnot\lnot$-closed subsets.  This is a significant restriction, which may
be validated in many cases of interests.  In contrast, the original hypothesis
(postulating minimal elements for ALL predicates)
is generally equivalent to every subset being decidable.
For example, the file \texttt{WFCounters.agda} shows that assuming the full minimality hypothesis
for the standard strict order on the natural numbers entails decidability
of every subset of $\nat$ --- which is clearly incompatible with effective semantics.

One could also consider proving well-foundedness ``up to $\lnot\lnot$'',
asserting that every element is $\lnot\lnot$-accessible,
every sequence $\lnot\lnot$-contains a non-decreasing index,
for every non-empty subset there $\lnot\lnot$-exists a minimal element, etc.
These are formalized in the definition below, and in the corresponding file
\texttt{WFWeakDefinitions.agda}. The logical relationships between all notions of well-foundedness are presented in Figure~\ref{fig:WF}.

@@

{
\def\arraystretch{1.3}
\begin{figure}[h!]
\small 
\begin{tabular}{@{}l l l @{}}
\toprule
\textbf{Property} & \textbf{Definition}  &\textbf{Agda Code}  \\
\midrule
$\rdec$   & $R$ is decidable 
          & $\forall$ \verb|{x} {y}| $\to$ \verb|EM (R x y)|\\
$\FB$     & $R$ is finitely branching 
          & $\mathtt{\forall (a : A) \to \Sigma [ xs \in List A ] (\forall b \to R a b \to b \in{}List\, xs)}$ \\
$\accCor$ & Accessibility is coreductive 
          & \verb|R -coreductive (R -accessible)| \\
$\accDNE$ & Accessibility is $\lnot\lnot$-closed
          & $\lnot\lnot$\verb|Closed (R -accessible)| \\
$\corDNE$ & Coreductives are $\lnot\lnot$-closed 
          & $\mathtt{\forall (P : \mathcal{P} A) \to R}$-$\mathtt{coreductive\ P \to \lnot\lnot Closed\ P}$\\ 
$\mpe$    & Sequences are $\lnot\lnot$-closed
          & $\forall\ \mathtt{(f : \nat \to A) \to \lnot\lnot Closed 
          (\lambda x \to \Sigma [ k \in \nat ] (f\, k \equiv x))}$ \\
\bottomrule
\end{tabular}
\centering
\caption{Classical Properties Used in the Well-Foundedness Diagram}
\label{tab:cprop}
\end{figure}
}

{
\def\arraystretch{1.3}
\begin{figure}[h!]
\small 
\begin{tabular}{@{}l l l @{}}
\toprule
\textbf{Property} & \textbf{Definition}  &\textbf{Agda Code}  \\
\midrule
$\rdec$   & $R$ is decidable 
          & $\forall$ \verb|{x} {y}| $\to$ \verb|EM (R x y)|\\
$\FB$     & $R$ is finitely branching 
          & $\mathtt{\forall (a : A) \to \Sigma [ xs \in List A ] (\forall b \to R a b \to b \in{}List\, xs)}$ \\
$\mpe$    & Sequences are $\lnot\lnot$-closed
          & $\forall\ \mathtt{(f : \nat \to A) \to \lnot\lnot Closed 
          (\lambda x \to \Sigma [ k \in \nat ] (f\, k \equiv x))}$ \\
$\corDNE$ & Coreductives are $\lnot\lnot$-closed 
          & $\mathtt{\forall (P : \mathcal{P} A) \to R}$-$\mathtt{coreductive\ P \to \lnot\lnot Closed\ P}$\\ 
$\accDNE$ & Accessibility is $\lnot\lnot$-closed
          & $\lnot\lnot$\verb|Closed (R -accessible)| \\
$\accCor$ & Accessibility is coreductive 
          & \verb|R -coreductive (R -accessible)| \\
\bottomrule
\end{tabular}
\centering
\caption{Classical Properties Used in the Well-Foundedness Diagram}
\label{tab:cprop}
\end{figure}
}
